---
title: "Supporting scientific research @ INBO"
author: "Dimitri Brosens, Peter Desmet, Daniel Du Seuil, Ivy Jansen, Els Lommelen, Jo Loos, Tanja Milotic, Thierry Onkelinx, Paul Quataert, Frederic Piesschaert, Hans Van Calster, Toon Van Daele, Floris Vanderhaeghe, Gert Van Spaendonck, Stijn Van Hoey, Pieter Verschelde, Toon Westra, Carine Wils"
date: "March 31, 2017"
output: 
  INBOmd::inbo_slides:
    location: Master of Statistical Data Analysis Day, Coupure Campus, UGent
    institute: Research Institute for Nature and Forest (INBO)
    theme: vlaanderen
    cover: "Image00012.png"
    cover_hoffset: "-30mm"
    toc: FALSE
---

```{r include = FALSE}
library(knitr)
opts_chunk$set(
  echo = FALSE
)
```

```{r}
staff <- 262
admin <- 31
wo <- c(BMK = 8, Data = 8, ICT = 6, Lab = 10, Bib = 8)
science <- staff - admin - sum(wo)
```

# INBO

## INBO

- Instituut voor Natuur- en Bosonderzoek / Research Institute for Nature and Forest
- Flemish research and knowledge centre for nature and its sustainable management and use
- conducts research and supplies knowledge to policy makers and stakeholders
- supports
    - agencies of Flemish government related with countryside or nature within urban areas
    - organisations involved in nature conservation, forestry, agriculture, hunting and fishery
- `r staff` staff
    - `r science` scientific staff (scientists + technicians)
    - `r sum(wo)` support for scientific staff
    - `r admin` management + administrative staff

## Data collection and provision

- INBO collects and provides data for international reporting
- Participates in (inter)national research networks
    - [LTER](https://lternet.edu): the Long Term Ecological Research network
    - [LifeWatch](http://www.lifewatch.eu): e-Science and Technology European Infrastructure for Biodiversity and Ecosystem Research 
    - [ALTER-Net](http://www.alter-net.info): A Long-Term Biodiversity, Ecosystem and Awareness Research Network
    - ...

## Support for scientific staff

- 5 specialised teams dedicated to support the scientific staff
    1. Biometry, Methodology & Quality assurance (`r wo["BMK"]`)
    1. Open data & open research (`r wo["Data"]`)
    1. Laboratory  (`r wo["Lab"]`)
    1. IT operations & development  (`r wo["ICT"]`)
    1. Library  (`r wo["Bib"]`)

# Biometry, Methodology & Quality assurance

## A scientific corporate identity and quality assurance

- involvement from design phase up to implementation phase
- creating, evaluating and adjusting procedures
    - harmonising protocols for data collection
    - validation of data through statistical modelling
- improving reproducibility
    - scripted analysis (R)
    - dynamic documents (RMarkdown)
    - version control (git)
- gradual change!
    - focus on early adopters and early majority
    - stimulation and support is vital
    - new procedures should lower the burden in the long run
    - allow for a long transition period before enforcing change

## Creating a methodological platform

- maintaining tools
    - software, manuals, documentation, books, ...
    - templates for basic analyses
    - corporate identity styles for graphics (ggplot2) and dynamics documents (markdown, LaTeX, HTML)
- implementing methodologies
    - standardise methods and analyses with focus on long term monitoring and common experiments
    - testing new methods and analysis techniques
    - developing new methods and analysis techniques
- improving self-reliance of scientists
    - internal statistics courses ranging from "stats 101" to generalised linear mixed models
    - offer scientists tools to help themselves and each other: websites, webinars, books, ...
    
## Leading specific projects

Some projects are led by the team, in collaboration with scientists. These projects typically:

- have more statistical or technological challenges
- strive towards achieving higher quality and standardization levels at INBO
- have a longer term

Examples:

- design and implementation of monitoring networks to support and evaluate Natura 2000 objectives
- design and implementation of monitoring of nature management plans
- elaboration of population modelling methods
- ... (add other projects ?) - please review

## Support and services

- teaching
    - statistical courses (sampling theory, GLM, GLMM, multivariate data analysis, ...)
    - R and RStudio
- consultancy
    - help desk: ad hoc questions on statistical techniques and R
    - sounding board: critical reflection on research goals, experimental design, statistical analysis, ...
- expertise assignments
    - experimental design: spatio-temporal resolution, sample size calculation, financial impact, ...
    - complex analyses requiring a full-fledged statistician
- quality control
    - scientists can ask for a review of their code
