---
title: "Supporting scientific research @ INBO"
author: "Dimitri Brosens, Peter Desmet, Daniel Duseuil, Ivy Jansen, Els Lommelen, Jo Loos, Tania Milotic, Thierry Onkelinx, Paul Quataert, Frederic Piesschaert, Hans Van Calster, Toon Van Daele, Floris Vanderhaeghe, Gert Van Spaendock, Stijn Van Hoey, Pieter Verschelde, Toon Westra, Carine Wils"
date: "March 31, 2017"
output: 
  INBOmd::inbo_slides:
    location: Master of Statistical Data Analysis Day, Coupure Campus, UGent
    institute: Research Institute for Nature and Forest (INBO)
    theme: vlaanderen
    cover: "Image00012.png"
    cover_hoffset: "-30mm"
    toc: FALSE
---

```{r include = FALSE}
library(knitr)
opts_chunk$set(
  echo = FALSE
)
```

```{r}
staff <- 262
admin <- 31
wo <- c(BMK = 8, Data = 8, ICT = 6, Lab = 10, Bib = 8)
science <- staff - admin - sum(wo)
```

# INBO

## INBO

- Instituut voor Natuur- en Bosonderzoek / Research Institute for Nature and Forest
- Flemish research and knowledge centre for nature and its sustainable management and use
- conducts research and supplies knowledge to policy makers and stakeholders
- supports
    - agencies of Flemish government related with countryside or nature within urban areas
    - organisations involved in nature conservation, forestry, agriculture, hunting and fishery
- `r staff` staff
    - `r science` scientific staff (scientists + technicians)
    - `r sum(wo)` scientific support staff
    - `r admin` management + administrative staff

## Data collection and provision

- INBO collects and provides data for international reporting
- Participates in (inter)national research networks
    - [LTER](https://lternet.edu): the Long Term Ecological Research network
    - [LifeWatch](http://www.lifewatch.eu): e-Science and Technology European Infrastructure for Biodiversity and Ecosystem Research 
    - [ALTER-Net](http://www.alter-net.info): A Long-Term Biodiversity, Ecosystem and Awareness Research Network
    - ...

## Scientific support staff

- 5 specialised teams dedicated to support the scientific staff
    1. Biometrics & Quality assurance (`r wo["BMK"]`)
    1. Open data & open research (`r wo["Data"]`)
    1. Laboratory  (`r wo["Lab"]`)
    1. IT operations & development  (`r wo["ICT"]`)
    1. Library  (`r wo["Bib"]`)

# Biometrics & Quality assurance

## A scientific corporate identity and quality assurance

- creating, evaluating and adjusting procedures
    - harmonising protocols for data collection
    - validation of data through statistical modelling
- improving reproducibility
    - scripted analysis (R)
    - dynamic documents (RMarkdown)
    - version control (git)
- gradual change!
    - focus on early adopters and early majority
    - stimulation and support is vital
    - new procedures should lower the burden in the long run
    - allow for a long transition period before enforcing change

## Creating a methodological platform

- maintaining tools
    - software, manuals, documentation, books, ...
    - templates for basic analyses
    - corporate identity styles for graphics (ggplot2) and dynamics documents (markdown, LaTeX, HTML)
- implementing methodologies
    - standardise analyses with focus on long term monitoring and common experiments
    - testing new analysis techniques
    - developing new analysis techniques
- improving self-reliance of scientists
    - internal statistics courses ranging from "stats 101" to generalised linear mixed models
    - offer scientists tools to help themselves and each other: websites, webinars, books, ...

## Support and services

- consultancy
    - help desk: ad hoc questions on statistical techniques and R
    - sounding board: critical reflection on research goals, experimental design, statistical analysis, ...
- expertise assignments
    - experimental design: spatio-temporal resolution, sample size calculation, financial impact, ...
    - complex analyses requiring a full-fledged statistician
- quality control
    - scientists can ask for a review of their code
