---
title: "Supporting scientific research @ INBO"
author: "Dimitri Brosens, Peter Desmet, Daniel Duseuil, Ivy Jansen, Els Lommelen, Jo Loos, Tania Milotic, Thierry Onkelinx, Paul Quataert, Frederic Piesschaert, Hans Van Calster, Toon Van Daele, Floris Vanderhaeghe, Gert Van Spaendock, Stijn Van Hoey, Pieter Verschelde, Toon Westra, Carine Wils"
date: "March 31, 2017"
output: 
  INBOmd::inbo_slides:
    location: Master of Statistical Data Analysis Day, Coupure Campus, UGent
    institute: Research Institute for Nature and Forest (INBO)
    theme: vlaanderen
    cover: "Image00012.png"
    cover_hoffset: "-30mm"
---

```{r include = FALSE}
library(knitr)
opts_chunk$set(
  echo = FALSE
)
```

```{r}
staff <- 262
admin <- 31
wo <- c(BMK = 8, Data = 8, ICT = 6, Lab = 10, Bib = 8)
science <- staff - admin - sum(wo)
```

# INBO

## INBO

- Instituut voor Natuur- en Bosonderzoek / Research Institute for Nature and Forest
- Flemish research and knowledge centre for nature and its sustainable management and use
- conducts research and supplies knowledge to policy makers and stakeholders
- supports
    - agencies of Flemish government related with countryside or nature within urban areas
    - organisations involved in nature conservation, forestry, agriculture, hunting and fishery
- `r staff` staff
    - `r science` scientific staff (scientists + technicians)
    - `r sum(wo)` scientific support staff
    - `r admin` management + administrative staff

## Data collection and provision

- INBO collects and provides data for international reporting
- Participates in (inter)national research networks
    - [LTER](https://lternet.edu): the Long Term Ecologisch Research network
    - [LifeWatch](http://www.lifewatch.eu): e-Science and Technology European Infrastructure for Biodiversity and Ecosystem Research 
    - [ALTER-Net](http://www.alter-net.info): A Long-Term Biodiversity, Ecosystem and Awareness Research Network
    - ...

## Scientific support staff

- 5 specialised teams dedicated to support the scientific staff
    1. Biometrics & Quality assurance (`r wo["BMK"]`)
    1. Open data & open research (`r wo["Data"]`)
    1. Laboratory  (`r wo["Lab"]`)
    1. IT operations & development  (`r wo["ICT"]`)
    1. Library  (`r wo["Bib"]`)
